---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# loop

**`{loop}`** contains several functions to perform loops in interesting ways.
The simplest function is `loop()`, which is built on by the other functions.

`loop` supports a dollar notation so that we can leverage the autompletion of
function arguments. Arguments to be looped on are prefixed with `+`.
Outputs are simplified by `type` if provided.

```{r}
library(loop)
loop(rep)(+11:12, 2)   # rather than `lapply(11:12, rep, 2)`
loop$rep(+11:12, 2)    # same thing, but we benefit from autocomplete
loop$rep(+11:12, +1:2) # rather than `Map(rep, 11:12, 1:2)`

# rather than `Map(function(...) rep(..., each =2), 11:12, 1:2)` :
loop$rep(+11:12, +1:2, each = 2) 
```

Outputs from `loop()` can be simplified by type
```{r}

l <- list(iris, cars)
loop(nrow, numeric)(+l) # rather than `vapply(l, nrow, numeric(1))`
```

The `type` can be any of the usual atomic types such as `double`, 
`numeric`, `character`, `logical`, plus any of four special tokens:

 * `df_rows` - returns a data.frame, where each result is a row
 * `mat_rows` - returns a matrix, where each result is a row
 * `mat_cols` - returns a matrix, where each result is a column
 * `simplify` - returns a simplified output, รก la `sapply()`

```{r}
x <- loop(rnorm)(1000,+1:10)
loop(mean,numeric)(+x)

loop(rnorm,mat_rows)(3,+1:4) # returns a 4x3 matrix
loop(rnorm,mat_cols)(3,+1:4) # returns a 3x4 matrix


```

To make use of both the type return and the dollar notation, we also
provide various wrapper functions around `loop(fun,type)` such as `loop_dbl()`,
`loop_int()` and `loop_chr()`.

```{r}
loop_dbl$mean(+x)
loop_int$length(+x)
loop_chr$paste0(+letters,+LETTERS)
```
When working with functions which take vectors as arguments, we would
normally wrap variables in `c()` to pass them as a single argument.
The parallel to this in `{loop}` is `c_loop()`

```{r}
x <- rnorm(10)
y <- rnorm(10)
z <- rnorm(10)
loop_dbl$mean(+c_loop(x,y,z))
```
This is particularly useful in `tidyverse` evaluations, and allows
for row-wise evaluation of commands.

```{r}
df <- data.frame(
  x = rnorm(3),
  y = rnorm(3),
  z = rnorm(3)
)

dplyr::mutate(df,m = loop_dbl$mean(+c_loop(x,y,z)))
```

A final example of where `c_loop()` can make iteration easier is in
a `for()` loop. We can iterate through the `c_loop()` elements in
parallel, by referring to them as elements of the iterator, either 
using `$` for named elements or their positional number

```{r}
for(. in c_loop(a=1:3,b=2:4)){
  print(.$a*.$b)
}

for(. in c_loop(1:3,3:5)){
  print(.[1]*.[2])
}

```



`looper()` also behaves like `for()` in that it takes an _expression_ 
provided between curly braces, `{` and `}`. Much like the above, it 
can iterate over several variables at once using the same notation 
as `loop()`, except we don't need to pre-pend variables with a `$`.
Much like `loop()`, it returns a variable, which can be simplified with
`type=`
```{r}
looper(
  seed = +c(100,300,500),
  n = +1:3, mean = +2:4,  sd = 1,
            lower = +0:2, upper=5,
  type = df_rows,
  {
    set.seed(seed)
    data.frame(
      col_1 = rnorm(n,mean,sd),
      col_2 = runif(n,lower,upper)
    )
  }
)
```


## Installation

Install with :

``` r
remotes::install_github("moodymudskipper/loop")
```

## Why ?

We get used to the apply functions, but : 

* They're hard to grasp at first
* No way to leverage the applied function's autocomplete
* The IDE highlights the apply function, not the applied function
* No need to shuffle the arguments when going from `lapply` to `Map` or `sapply` to `mapply`
* Much more compact to use constant arguments with `Map` and `sapply`
* I believe its more intuitive to have the applied function first and distinct from the arguments
